Machine learning topological phases in real space

### 2020-06-25 8_prb_submission/1_reply_referee/article_with_comments/article.tex
Further comments

\vspace{0.25cm}
\textcolor{blue}{\citequote{pipeline_not_work} We can think of several use cases for the eigenvector ensembling procedure outlined in the paper. Among these, we emphasize its applicability in tackling the so-called "curse of dimensionality" that plagues many problems in condensed matter and statistical physics. In the paper, we have demonstrated an application in which the dimensionality of an SSH lattice can be greatly reduced in so far as its topological properties are concerned, simply by extracting the most relevant lattice sites in its information entropy signature. Similar feature extraction and dimensionality reduction procedures are often employed in machine learning applications and are surely to become the norm in data-driven physics. The eigenvector ensembling algorithm can thus be viewed as a general framework in this direction. Similarly, by thoughtfully attaching a probability distribution to the eigenvectors, we end up with a powerful sampling technique that may ultimately lead to great reduction in the dimensionality of a Monte Carlo simulation, while still capturing essentially all the relevant physics pertaining to the problem. We have added a paragraph to the \textbf{Discussion} section outlining these possibilities. As for the novelty of the method, the eigenvector ensembling algorithm together with the explainability powers of decision trees allowed us to rigorously investigate the information entropy of topological phase transitions. To our knowledge, our work pioneers the discovery of new concepts in physics through model explainability in machine learning.  
}

\citequote{sufficient_accuracy} ...

\citequote{superior_comparable} Mention trends in model explainability as well as performance scores.

@@

\citequote{unsupervised_approaches} Mention Hamiltonian compression, of unsupervised learning through dimensional reduction.

@@

\textcolor{blue}{As of yet model explainability is one of the topics at the forefront of machine learning research that has been little explored by the physics community working at the interface between the two disciplines}\citequote{unsupervised_learning}. The Shannon information entropy signals and the corresponding entropic uncertainty relations associated with topological phase transitions presented for the first time in this work provide a clear illustration of how model explainability in machine learning can guide new discoveries in condensed matter and quantum materials physics, since the existence of these signals was established by analyzing data from several thousand SSH systems which, taken individually, could not have provided any concrete hint of their existence.\citequote{pipeline_not_work}

\begin{figure}
\centering
\subfigure[]{\label{ssh1}\includegraphics[width=.45\textwidth]{./phase_diagrams/ssh1.png}}\quad
\subfigure[]{\label{ssh2}\includegraphics[width=.45\textwidth]{./phase_diagrams/ssh2.png}}
\caption{Phase diagrams in parameter space. a) SSH model with first-neighbor hoppings $t_1$ and $t_2$. The (red) regions with winding number $W$ = 0 are trivial, while the (blue) regions with winding number $W$ = 1 are topologically non-trivial. b) SSH model with first ($t_1$ and $t_2$) and second ($T_1$ and $T_2$) nearest-neighbor hoppings. In this article we set $t_1$ = $t_2$ = 1 and renamed the variables $T_1$ $\rightarrow$ $t_1$, $T_2$ $\rightarrow$ $t_2$ for convenience. The (orange) region with winding number $W$ = 0 is trivial while the others with winding numbers $W$ = -1, $W$ = 1 and $W$ = 2 (red, green and blue respectively) are topologically non-trivial.}
\label{fig:phasediagrams}
\end{figure}

@@

\textcolor{blue}{As performance metrics, we report here both accuracy of predicted classes for eigenvectors as well as accuracy of predicted classes for Hamiltonians obtained from eigenvector ensembling. These accuracy scores are to be gauged against the baseline of a system that simply guesses the most frequent class for all eigenvectors (or Hamiltonians). Checking against this baseline is important because it indicates whether the decision trees are in fact learning the underlying patterns that relate real space coordinates to winding numbers, and therefore whether the associated information entropy signature is meaningful or not.}\citequote{sufficient_accuracy}

@@

>>>>

\vspace{0.25cm}
\textcolor{blue}{\citequote{pipeline_not_work} We can think of several use cases for the eigenvector ensembling procedure outlined in the paper. Among these, we emphasize its applicability in tackling the so-called "curse of dimensionality" that plagues many problems in condensed matter and statistical physics. In the paper, we have demonstrated an application in which the dimensionality of an SSH lattice can be greatly reduced in so far as its topological properties are concerned, simply by extracting the most relevant lattice sites in its information entropy signature. Similar feature extraction and dimensionality reduction procedures are often employed in machine learning applications and are surely to become the norm in data-driven physics. The eigenvector ensembling algorithm can thus be viewed as a general framework in this direction. Similarly, by thoughtfully attaching a probability distribution to the eigenvectors, we end up with a powerful sampling technique that may ultimately lead to great reduction in the dimensionality of a Monte Carlo simulation, while still capturing essentially all the relevant physics pertaining to the problem. We have added a paragraph to the section \textbf{Discussion} outlining these possibilities. As for the novelty of the method, the article binds together contemporary concepts in physics (topological states of matter and information theory) to a leading research theme in machine learning (model explainability). To our knowledge, our work pioneers the discovery of new concepts in physics through model explainability in machine learning. We have added a last paragraph to the introduction emphasizing this point.  
}

\textcolor{blue}{\citequote{sufficient_accuracy}, \citequote{superior_comparable} We use the term "sufficient accuracy" to denote a system that performs considerably better at predicting the phases of Hamiltonians than the baseline system that simply guesses the most frequently occurring class for all Hamiltonians. This gauge is needed because it ensures that the resulting entropy signatures are meaningful in the sense that they encode a summary of where useful information about the topological phase of a system can be found. We have added a paragraph to the section \textbf{Numerical experiments} to emphasize the use of this baseline. As to "high accuracy", we note that the accuracy scores obtained by us are comparable to the highest scores reported in Phys. Rev. Lett. 120, 066401, where neural networks were trained on wavevector space data in a similar supervised learning task. This second gauge further reinforces that the entropy signatures presented in the article indeed express a realistic summary of the availability of topological information from each lattice site. A small paragraph was added to the section \textbf{Discussion} to explicitly note this point.}

@@

@@

The Shannon information entropy signals and the corresponding entropic uncertainty relations associated with topological phase transitions presented for the first time in this work provide a clear illustration of how model explainability in machine learning can guide new discoveries in condensed matter and quantum materials physics, since the existence of these signals was established by analyzing data from several thousand SSH systems which, taken individually, could not have provided any concrete hint of their existence.\citequote{pipeline_not_work}

\textcolor{blue}{As of yet model explainability is one of the topics at the forefront of machine learning research that has been little explored by the physics community working at the interface between the two disciplines. By tying together contemporary concepts in physics (topological phase transitions and information theory) to a topic at the edge of machine learning research (model explainability) we hope to append the latter to the growing list of machine learning tools at physcists' disposal.  }\citequote{pipeline_not_work}
\begin{figure}
\centering
\subfigure[]{\label{ssh1}\includegraphics[width=.45\textwidth]{./phase_diagrams/ssh1.png}}\quad
\subfigure[]{\label{ssh2}\includegraphics[width=.45\textwidth]{./phase_diagrams/ssh2.png}}
\caption{Phase diagrams in parameter space. a) SSH model with first-neighbor hoppings $t_1$ and $t_2$. The (red) regions with winding number $W$ = 0 are trivial, while the (blue) regions with winding number $W$ = 1 are topologically non-trivial. b) SSH model with first ($t_1$ and $t_2$) and second ($T_1$ and $T_2$) nearest-neighbor hoppings. In this article we set $t_1$ = $t_2$ = 1 and renamed the variables $T_1$ $\rightarrow$ $t_1$, $T_2$ $\rightarrow$ $t_2$ for convenience. The (orange) region with winding number $W$ = 0 is trivial while the others with winding numbers $W$ = -1, $W$ = 1 and $W$ = 2 (red, green and blue respectively) are topologically non-trivial.}
\label{fig:phasediagrams}
\end{figure}

@@

\textcolor{blue}{As performance metrics, we report here both accuracy of predicted classes for eigenvectors as well as accuracy of predicted classes for Hamiltonians obtained from eigenvector ensembling. These accuracy scores are to be gauged against the baseline of a system that simply guesses the most frequent class for all eigenvectors (or Hamiltonians). Checking against this baseline is important because it indicates whether the decision trees are in fact learning the underlying patterns that relate real space coordinates to winding numbers, and therefore whether the associated information entropy signature is meaningful or not.}\citequote{sufficient_accuracy} \citequote{evaluation_metrics}

@@

\textcolor{blue}{The accuracy scores obtained in this paper were comparable to those reported in \cite{PhysRevLett.120.066401}, where dense and convolutional neural networks were trained on wavevector space data to predict the winding numbers of SSH Hamiltonians via supervised learning. This high accuracy level serves as a strong evidence that the entropy signatures presented here indeed express where topological information can most readily be recovered from in the SSH lattices investigated.}\citequote{superior_comparable}
